---
title: "Class 6 - Homework"
author: "Deborah Sigaud Kutner (PID: A69027018)"
date: 01/26/2024
format: pdf
---
## Instructions

**Q6. How would you generalize the original code above to work with any set of input protein structures?**

library(bio3d) 

s1 <- read.pdb("4AKE")  # kinase with drug 

s2 <- read.pdb("1AKE")  # kinase no drug 

s3 <- read.pdb("1E4Y")  # kinase with drug 


s1.chainA <- trim.pdb(s1, chain="A", elety="CA") 

s2.chainA <- trim.pdb(s2, chain="A", elety="CA") 

s3.chainA <- trim.pdb(s3, chain="A", elety="CA") 


s1.b <- s1.chainA$atom$b 

s2.b <- s2.chainA$atom$b 

s3.b <- s3.chainA$atom$b 


plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor") 

plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor") 

plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")


## Function "protein":

- "code" is the main argument of the function, requiring the PDB filename or the four letter PDB identifier. 
- "chain", "elety" and "typ" have defaults, but can be changed.

```{r , warning = FALSE}

#install.packages('bio3d') # install package
library(bio3d) # load library

protein <- function(code, chain = "A", elety="CA", typ="l") {  
  pdb <- read.pdb(code)                                    
  # reads the pdb code for the desired protein
  chain <- trim.pdb(pdb, chain=chain, elety=elety)         
  # loads the chain information
  plotb3(chain$atom$b, sse=chain, typ=typ, ylab="Bfactor", main = paste("PDB:", code)) 
  # returns the plot of B-factor values and PDB residue numbers. 
  # the type of plot is changeable (e.g. line, points, etc.).
  }

protein(code = "4AKE") # check if function is working for one code

```

Loading multiple proteins at the same time using the apply function.

```{r , warning= FALSE}

files <- matrix(c("4AKE", "1AKE", "1E4Y"))
apply((files), 1, protein)
```


